Fast gravitational wave parameter estimation without compromises

### 2023-07-25 src/tex/ms.tex
Remove space that block compilation
Add citation to flowMCv0.1

\corr{The interplay between \textsc{flowMC} and GPU acceleration is critical to
the performance of the code. A standard PE run using a nested sampler or
standard MCMC requires around $10^6$ waveform evaluations, Assuming $10^5$
frequency bins per waveform, that corresponds to evaluating the waveform model
at $10^11$ frequency-parameter points. One would think this can be achieved
within $\sim 100\textrm{s}$ by a A100 GPU given the benchmark mentioned above.
However, a sampler requires sequential evaluation to accept or reject a sample,
where what a GPU provide is parallel evaluation of the waveform, i.e. it can run
many chains in parallel, but not many steps with the same amount of computation.
This means one cannot just obtain the posterior samples by running the waveform
model on a GPU with a typical sampler, in the sense that we cannot run many
chains for one step and expect unbias posterior samples.

This is precisely why \textsc{flowMC} is necessary for our approach. Since every
accepted sample generated using the normalizing flow as proposal is nearly independent,
and the sampling from the normalizing flow is an embarrassingly parallel
process, the name of the game then becomes training the normalizing flow as fast as
possible. Having many chains helps to produce the necessary amount of samples
for training the normalizing flow, which means now we can meaningfully harvest
the benefit of having large throughput of GPU. Without either of this two,
our approach would be significantly less performant.}

@@

\section{Acknowledgements}
We thank Will M.~Farr, Aaron Zimmerman, Daniel Foreman-Mackey and Marylou Gabri\'e for helpful discussions; we also thank Carl-Johan Haster, Neil J.~Cornish and Thomas Dent for comments on the draft.
The Flatiron Institute is a division of the Simons Foundation.
This material is based upon work supported by NSF's LIGO Laboratory which is a major facility fully funded by the National Science Foundation.
This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA. LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector. Additional support for Advanced LIGO was provided by the Australian Research Council. Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain. The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT), and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea, Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
T.E.\ is supported by the Horizon Postdoctoral Fellowship.
\software{scipy \citep{2020SciPy-NMeth}, flowMC \citep{2022arXiv221106397W,Gabrie:2021tlu}, ripple package \citep{Edwards:2023sak}, pBilby \citep{Smith:2019ucc}}

>>>>

\corr{The interplay between \textsc{flowMC} and GPU acceleration is critical to
the performance of the code. A standard PE run using a nested sampler or
standard MCMC requires around $10^6$ waveform evaluations, Assuming $10^5$
frequency bins per waveform, that corresponds to evaluating the waveform model
at $10^11$ frequency-parameter points. One would think this can be achieved
within $\sim 100\textrm{s}$ by a A100 GPU given the benchmark mentioned above.
However, a sampler requires sequential evaluation to accept or reject a sample,
where what a GPU provide is parallel evaluation of the waveform, i.e. it can run
many chains in parallel, but not many steps with the same amount of computation.
This means one cannot just obtain the posterior samples by running the waveform
model on a GPU with a typical sampler, in the sense that we cannot run many
chains for one step and expect unbias posterior samples.
This is precisely why \textsc{flowMC} is necessary for our approach. Since every
accepted sample generated using the normalizing flow as proposal is nearly independent,
and the sampling from the normalizing flow is an embarrassingly parallel
process, the name of the game then becomes training the normalizing flow as fast as
possible. Having many chains helps to produce the necessary amount of samples
for training the normalizing flow, which means now we can meaningfully harvest
the benefit of having large throughput of GPU. Without either of this two,
our approach would be significantly less performant.}

@@

\section{Acknowledgements}
We thank Will M.~Farr, Aaron Zimmerman, Daniel Foreman-Mackey and Marylou Gabri\'e for helpful discussions; we also thank Carl-Johan Haster, Neil J.~Cornish and Thomas Dent for comments on the draft.
The Flatiron Institute is a division of the Simons Foundation.
This material is based upon work supported by NSF's LIGO Laboratory which is a major facility fully funded by the National Science Foundation.
This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA. LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector. Additional support for Advanced LIGO was provided by the Australian Research Council. Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain. The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT), and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea, Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
T.E.\ is supported by the Horizon Postdoctoral Fellowship.
\software{scipy \citep{2020SciPy-NMeth}, flowMC \citep{2022arXiv221106397W,Gabrie:2021tlu, kaze_wong_2023_7706605}, ripple package \citep{Edwards:2023sak}, pBilby \citep{Smith:2019ucc}}
